What's hot and what's not:\\Windowing developer topic analysis

### 2009-04-11 lda-paper.tex
Getting closer to a 10 pager + migod edits

Topic analysis, with respect to Software Control Systems (SCSs) such
as CVS, Subversion, Git etc., is potentially valuable as it has many
uses for various stakeholders and uses.  Imagine a development team
has agreed to implement some features at the start of a development
iteration. By the end of the iteration only some of the agreed upon
features are implemented. The developers insist they were focusing
 on the features agreed upon, but what actually happened? Topic
analysis could help by teasing out the distinct topics or focusses of
the developers. Perhaps the developers had to deal with other pressing issues.
Developers might have grouped those orthogonal issues into the
features they were working on, thus when asked about how their time
was spent the orthogonal changes would not be covered. This is the
kind of underlying story that we hope that windowed topic analysis can
uncover. If windowed topic analysis can discover important but
overlooked topics, it can help both developers and managers. Developers
can see a summary of what worked on, and managers can development.

Commits in SCSs are not necessarily consistent and they can have
multiple purposes. Thus we hope that our research will help uncover
these multiple purposes that manifest themselves as topics interleaved
within the commits.  Others have demonstrated that by using 
topic analysis techniques, such as Latent Semantic Indexing (LSI) and
Latent Dirichlet Allocation (LDA), can aid in separating topics from
these mixed comments. These topics often describe distinct ideas that
are interleaved throughout the development. A topic represents a
groups of commit comments that are related to each other by their
content.  In this case a topic is a set of tokens.

@@

One use of topic analysis is to help us partition a project's
time-line into periods of developer focus. By breaking apart an
iteration into sets of topics and trends (topics which recur), we may
be able to recognize the underlying software development process from
these commits. Alternatively, we can determine what particular
maintenance task was being performed are at a given time.

We propose to extend the use of common existing topic analysis techniques by
applying them to windows of documents over time. We hope by looking
at topic clusters during periods or windows over time we can identify
ongoing areas of focus on a select group of topics as well reflect more intermittent topics. Figure \ref{fig:lda} illustrates
the kind of visualization we want to automatically generate. We want
to see topics that are unique to a certain time period, displayed
across the time axis. Those topics that recur, or occur over a
larger period are plotted continuously. In our example we have
titled each topic with a single word.

@@

We now define some terms that we will use in the rest of the paper: A
\emph{message} is a block of text written by developers. In this
paper, messages will be the CVS commit comments made when the user
commits changes to files in a repository. A \emph{word distribution}
is the summary of a message by word count. Each word distribution will
be over the words in all messages. However, most words will not appear
in a message, a word distribution is effectively a word count divided
by the message size. A \emph{topic} is a word distribution. It is
essentially a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is a one or more similar topics which recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project.

@@

In terms of clustering and finding topic distributions, Latent
Dirichlet Allocation (LDA)~\cite{944937} competes with Latent Semantic
Indexing
(LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17},
probabilistic Latent Semantic Indexing (pLSI)~\cite{944937} and
semantic clustering~\cite{1698774,1566153}. These tools are used for
document modelling, document clustering and collaborative
filtering. LDA attempts to encode documents as mixture
model, a combination, of topics.  LDA is used in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
is to extract topics from documents such as methods, bug
reports, source code and files.

LDA, LSI and semantic clustering extract topic
clusters from the documents that are independent from one another. LDA
does not name these clusters but the clusters consist of words whose
relationship is often obvious to someone familiar with the corpus. We
could swap LDA for LSI or semantic clustering and produce similar
results. Our data is posed as documents with word distributions (word
counts per documents) and LDA extracts distinct topics (clusters of
words) from the documents.

@@

In order to infer or associate the expertise of an author with topics
extracted from SCS, Linstead et al. proposed an author-source-code model
using LDA~\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}. This model
is essentially an extension of the author-topic (AT)
model~\cite{1036902}, which associated authors and topics extracted
with LDA.

@@

LSI is related to LDA and has been used to identify topics in software
artifacts for formal concept analysis and concept
location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
Concept analysis aims to automatically extract concepts from source
code or documents that are related to the source code.  Concept location concerns how
high level concepts relate to low level entities such as source code. For
example, when fixing a bug how and where do the concepts in the bug
report map back to the source code?  Semantic Clustering has also been
used for similar purposes~\cite{1698774,1566153} as it is similar to
LSI.

@@

Our technique differs from the previous techniques because we apply
LDA locally to month-long windows of commit comments, where-as other
approaches apply LDA once to the entire project. This allows us to
examine the time-based windowing of topics over its development
history.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.9\textwidth]{lda}
  \caption{Example of topics extracted from MySQL 3.23. This is the
    kind of plot we eventually want to illustrate: named topics and
    topic trends. The horizontal axis is time by month examined. The
    vertical axis is used to stack topics that occur at the same
    time. Longer topics are topics which recur in adjacent
    window. Colors are arbitrary.}
  \label{fig:lda}
\end{figure*}

\shrinkit
\section{Methodology}
\shrinkit
\begin{figure*}
  \centering
  \includegraphics[width=0.9\textwidth]{commit-to-topics} 
  \caption{How commits are analyzed and aggregated into Topics and Trends. Commits are first extracted, then abstracted into word counts or word distributions. These word distributions are then given a topic analysis tool like LDA. LDA finds independent word distributions (topics) that these documents are related to.}
  \label{fig:commits}
\end{figure*}

@@

The \emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to
sink the larger trends to the bottom.  Once the larger trends are
stacked we fill in the gaps with the smaller trends in the same
window, and then stack the smaller trends on top.  Although there is a
chance that gaps will be left due to non-optimal stacking in practice
there are lot of small trends (90\% of all trends contain 1 topic)
which fill in these gaps quickly.  Different instances of the same
trend share the same color; apart from that, the color of trends is
randomly chosen and color similarity is not meaningful, this is easier
to see in the zoomed view of MaxDB 7.500's compact-trend-view shown in
Figure \ref{fig:zoomedsmear}.  The compact-trend-view is convenient
because it shows all of the information and the repeating continuous
trends are easy to pick out, although repeating discontinuous trends
are harder to spot.

@@

The {trend-histogram} (Figure \ref{fig:histogram}) superficially
resembles the trend-time-line.  However, in this view the trends are
plotted together by stacking to the left of their row, thus time
information is lost.  The trends are ordered by the number of topics
in the trend.  The trend-histogram shows the count of instances of a
trend and thus indicates which trends occur the most. Unfortunately
due to the large number of topics given allotted space it is often
best to crop off the trends with only one topic, otherwise the tail is
long.

%XXX Describe something useful like a trend
\begin{figure*}
  \centering
  \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
  \caption{Compact-trend-view shows topics per month for MaxDB
    7.500. The x-axis is time in months, the Y axis is used to stack
    topics occurring at the same time. Trends that are continuous are
    plotted as continuous blocks. The top 10 words in the topics are
    joined and embedded in box representing the topics.}
  \label{fig:topicsmear}
\end{figure*}

@@

\shrinkit
\subsection{Datasets and Tools}
\shrinkit

To extract data from CVS and BitKeeper we used softChange to extract
CVS repositories and provide a schema for our revision data. We used
bt2csv to extract BitKeeper revisions from BitKeeper web repositories.

We extracted the repositories of: PostgreSQL, MaxDB and
Firebird.

Our analysis tools consisted of: Hiraldo-Grok, an OCaml-based variant
of the Grok query language; Gnuplot, a graph plotting package; lda-c,
a LDA package implemented by Blei et al. \cite{944937}; and our topic
plotter, implemented in Haskell.

@@

@@

When we examined PostgreSQL, we did not find many trends with two or
more topics with a similarity of $7/10$. These trends were not very
large, lasting on 3 months at most. The first, second largest trends
were dedicated to the work of two external developers: Dal Zotto, Dan
McGuirk. The fifth largest trend related to work by PostgreSQL developer D'Arcy. Other
topics of the larger trends were changes to the TODO list in HTML, and
time string formating topics relating to timezones.

If we kept the stop words we found that the large trend consisted
mostly of stop words and non-stop words such as \emph{patch},
\emph{fix}, \emph{update}.  By decreasing the similarity constraint to
$1/2$ similarity the largest most the largest most common trend, which
stretched across development, contained these same words
(\emph{patch}, \emph{fix}, \emph{update}). The second most common
topic mentions Dal Zotto, while the third most common topic mentions
the \texttt{[PATCHES]} mailinglist and the names of some patch
contributors.  Other repeating topics refer to portability with Win32,
Central Europe Time (CEST) from email headers, issues with ALTER TABLE
and CVS branch merging (because CVS doesn't record merges explicitly).

@@

%XXX Edit
% Was reading lda.report
Bug tracker URLs dominated unique topics during some months. For
instance in the last month of MaxDB 7.500 development every topic
contained 1 unique Bug tracker URL, this pattern did not occur in the
previous month. We investigated the actual revisions and we did find
there was difference in behavior between and that the last month of
development commits were mostly referencing the bug tracker. If the
topics of one month were 20 or more unique bug tickets being
addressed, the global LDA analysis would probably miss this, yet these
bug tickets were the focus of development for that month and not
necessarily globally relevant.

The query optimizer was a topic which recurred during MaxDB's
development. In our plots, topics which mention optimizer occur four
times yet in the global-trend-view (Figure \ref{fig:statictopics}) it
is not in any of the topics. A query optimizer is rather important to
an DBMS, but as we have shown it just does not pop up as a topic on
its own. We tried to remove words to see if we could get an topic
optimizer topic. After removing stop words and then two of the most
common words the global analysis finally had a topic with optimizer in
its topic ten words. This shows that optimizer was important but it
had been obscured, while it would have been noticed using a more
locally relevant approach that we propose.

We noticed that commits which mentioned \emph{PHP} occurred two thirds
less frequently than commits which mentioned \emph{Perl}, but
\emph{Perl} related topics appeared in the global static topics for
MaxDB while \emph{PHP} related topics did not.  Our local topic
analysis mentioned \emph{PHP} in 5 different topics, yet only
mentioned \emph{Perl} in 4 different topics. Perhaps the Perl mentions
were not as independent topics as the PHP topics were.

We noticed that Perforce and Changelist were in just about every
topic, so we added them to the stop words list. As a result, longer
trends were shortened and in some cases the total number of topics
found per month was less.  Evaluating different similarity thresholds
showed that by removing common words one reduces the general
similarity of topics. That said the larger topics still existed. Thus
if more relevant stop words are added one should tune your topic
similarity parameters to handle these changes.

@@

Previous work on topic analysis that employed LSI and LDA typically
extracted a certain number of topics and tracked them over time. This
means that a single, static set of topics was tracked across the whole
development history of a project. \emph{Global topic analysis} is
topic analysis applied to entire lifetime of a project.

We carried out the same kind of analysis, as shown in
Figure~\ref{fig:statictopics}. We extracted 20 topics and plotted the
number of messages per month that were related to that topic. We found
was that one topic would often dominate the results, as shown in the
third row of Figure \ref{fig:statictopics}, while messages related to
other topics did not appear very often.
% This technique seems useful if the 20 extracted topics are relevant
% and useful, that said one weakness of this approach is that documents
% for a topic might only occur in one time window, yet we are tracking
% messages that could be related to it over the entire lifetime of the
% project.
This approach seems reasonable if most of the extracted topics are of
broad interest during most of the development process.  However, it
may be that some topics are of strong interest but only briefly; in
such a case, a window-based model gives a much stronger indication of
the fleeting importance of such topics, and can help to put such a
topic into its proper context.

@@

If we approach the difference of global topic analysis and
windowed topic analysis via common tokens we can see that common token
tend to dominate both results, at least in the case of MaxDB 7.500 our
local topic approach netted us topics which contained these relatively
important and common words: \emph{UNICODE}, \emph{DEC/OSF1},
\emph{ODBC}, \emph{crash}, \emph{SQLCLI}, \emph{SYSDD.cpnix},
\emph{backup}, \emph{select}, \emph{make}, \emph{memory}, \emph{view},
and finally \emph{debug}. None of these tokens were part of the global 
topic analysis topics, but they were part of 566 commits (6\% of the
entire system) to MaxDB 7.500. These tokens were part of 87 out of 520
(26 months, 20 topics per month) of our locally generated topics.

@@

Even with our liberal topic similarity metrics that produced both
long and short trends, we showed that there are only a few trends in a
repository which recur. Since so few trends recur and so many
trends appear only once this suggests that global topic analysis might
be ignoring locally unique topics.

The utility of global topic analysis is questionable if the value of
information decreases as it becomes older. Perhaps older trends will
dominate the topic analysis. Windowed localized topic analysis show
what are the unique topics.

\begin{figure}
  \centering
  %\includegraphics[width=0.45\textwidth]{maxdb7500-everything-by-month}
  \includegraphics[width=0.46\textwidth]{month}
  \caption{MaxDB 7.500 topics analyzed with global topic analysis, 20
    topics (each row) and their document counts plotted over the
    entire development history of MaxDB 7.500 (26 months). The shade
    of the topic indicates the number of documents matching that topic
    in that month relative to the number of documents (white is most,
    black is least).}
  \label{fig:statictopics}
\end{figure}

@@

In this study we are explicitly trusting that the programmers annotate
their changes with relevant information. We rely on the descriptions
they provide. If the language of check-in comments was automated we
would only be analyzing that.

We truncated the topics to top 10 tokens, this summary of a topic
might not have been as useful as determining the actual topic distance
between two word topic distributions. By truncating tokens we could be
losing valuable data.

Each project seemed to have their own kind of stop words, we did not
remove them, but perhaps dropping some of these stop words would aid
the clarity of topics produced through topic analysis; alternatively
it might cause our topic similarity plots to be fundamentally
different.

@@

We proposed and demonstrated the application of windowed topic
analysis, that is, which is topic analysis applied to commit messages during
periods of development. This approach differs from previous work that
would apply topic analysis globally to a project. We showed that many
topics that exist locally are relevant and interesting yet would often
not be detected via global topic analysis. We identify recurring topics
with a topic similarity measure which looks for topics which recur and
mutate repeatedly throughout the development of the software project.

The value of windowed topic analysis was demonstrated on a case study
of three database systems with a particular focus on MaxDB 7.500. We
showed that global topic analysis missed important topics like ODBC
changes.

@@

\bibliographystyle{latex8}
\bibliography{lda-paper}

>>>>

% XXX Come back here, clear up and cut 
% Topic analysis, with respect to
% Software Control Systems (SCSs) such as CVS, Subversion, Git etc.,
% is potentially valuable as it has many uses for various
% stakeholders.  Imagine a development team has agreed to implement
% some features at the start of a development iteration.  By the end
% of the iteration only some of the agreed upon features are
% implemented.  The developers insist they were focusing on the
% features agreed upon, but what actually happened?  Topic analysis
% could help by teasing out the distinct topics or focusses of the
% developers.  Perhaps the developers had to deal with other pressing
% issues.  Developers might have grouped those orthogonal issues into
% the features they were working on, thus when asked about how their
% time was spent the orthogonal changes would not be covered.  This is
% the kind of underlying story that we hope that windowed topic
% analysis can uncover. If windowed topic analysis can discover
% important but overlooked topics, it can help both developers and
% managers.  Developers and Managers can see a summary of what worked
% on and the topics of development.

Imagine that a development team have agreed upon a set of features to
implement for their next interation, but these features were not
finished by the end of the iteration. The developers insist they were
focussed on the features, yet their manager and is not sure what
happened. In a scenario like this, topic analysis can help tease out
the topics of development that the developers were focussed on. Topic
analysis could highlight the orthogonal issues that developers were
spending time on.  Topic analysis extracts groups of words which from
documents (commit log messages) which can be used to relate
documents. These extracted topics often correlate with actual
development topics that the developers discussed. By applying topic
analysis locally to an iteration, both the developers and managers
might be able to see what were the topics of development.

Commits in SCSs can have
multiple purposes. Thus we hope that our research will help uncover
these multiple purposes that manifest themselves as topics interleaved
within the commits.  Others have demonstrated that 
topic analysis techniques, such as Latent Semantic Indexing (LSI) and
Latent Dirichlet Allocation (LDA), can aid in separating topics from
these mixed comments. These topics often describe distinct ideas that
are interleaved throughout the development. A topic represents a
groups of commit comments that are related to each other by their
content.  In this case a topic is a set of tokens.

@@

Topic analysis can aid in partitioning a project's
time-line into periods of developer focus. By breaking apart an
iteration into sets of topics and trends (topics that recur), we may
be able to recognize the underlying software development process from
these commits. Alternatively, we can determine what particular
maintenance task was being performed at a given time.

We propose to extend existing topic analysis techniques by
applying them to windows of documents over time. We hope by looking
at topic clusters during periods or windows over time we can identify
ongoing areas of focus on a select group of topics as well reflect more intermittent topics. Figure \ref{fig:lda} illustrates
the kind of visualization we want to automatically generate. We want
to see topics that are unique to a certain time period, displayed
across the time axis. Those topics that recur, or occur over a
larger period are plotted continuously. In our example we have
titled each topic with a single word.

@@

We now define some terms that we will use in the rest of the paper: A
\emph{message} is a block of text written by developers. In this
paper, messages will be the CVS commit comments made when the user
commits changes to files in a repository. A \emph{word distribution}
is the summary of a message by word count. Each word distribution will
be over the words in all messages. However, most words will not appear
in a message, a word distribution is effectively a word count divided
by the message size. A \emph{topic} is a word distribution. It is
essentially a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is a one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project.

@@

In terms of clustering and finding topic distributions, Latent
Dirichlet Allocation (LDA)~\cite{944937} competes with Latent Semantic
Indexing
%(LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17},
(LSI)~\cite{1374321,10.1109/ICPC.2007.13},
probabilistic Latent Semantic Indexing (pLSI)~\cite{944937} and
semantic clustering~\cite{1698774,1566153}. These tools are used for
document modelling, document clustering and collaborative
filtering. LDA attempts to encode documents as a mixture
model, a combination of topics.  LDA is used in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20}%,NIPS2007637,1321709}
 to extract topics from documents such as methods, bug
reports, source code and files.

LDA, LSI and semantic clustering extract topic
clusters from the documents that are independent from one another. LDA
does not name these clusters but the clusters consist of words whose
relationship is often obvious to someone familiar with the corpus. We
could swap LDA for LSI or semantic clustering and likely produce similar
results. Our data is posed as documents with word distributions (word
counts per documents) and LDA extracts distinct topics (clusters of
words) from the documents.

@@

In order to infer or associate the expertise of an author with topics
extracted from SCS, Linstead et al. proposed an author-source-code model
using LDA~\cite{10.1109/MSR.2007.20}%,NIPS2007637,1321709}. 
This model
is essentially an extension of the author-topic (AT)
model~\cite{1036902}, which associated authors and topics extracted
with LDA.

@@

LSI is related to LDA and has been used to identify topics in software
artifacts for formal concept analysis and concept
%location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
location~\cite{1374321,10.1109/ICPC.2007.13}
Concept analysis aims to automatically extract concepts from source
code or documents that are related to the source code.  Concept location concerns how
high-level concepts relate to low level entities such as source code. For
example, when fixing a bug how and where do the concepts in the bug
report map back to the source code?  Semantic Clustering has also been
used for similar purposes~\cite{1698774,1566153} as it is similar to
LSI.

@@

Our technique differs from the previous techniques because we apply
LDA locally to month-long windows of commit comments, whereas other
approaches apply LDA once to the entire project. This allows us to
examine the time-based windowing of topics over its development
history.

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{lda}
  \caption{Example of topics extracted from MySQL 3.23. This is the
    kind of plot we eventually want to illustrate: named topics and
    topic trends. The horizontal axis is time by month examined. The
    vertical axis is used to stack topics that occur at the same
    time. Longer topics are topics which recur in adjacent
    window. Colors are arbitrary.}
  \label{fig:lda}
\end{figure*}

\shrinkit
\section{Methodology}
\shrinkit
\begin{figure*}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{commit-to-topics} 
  \caption{How commits are analyzed and aggregated into Topics and Trends. Commits are first extracted, then abstracted into word counts or word distributions. These word distributions are then given a topic analysis tool like LDA. LDA finds independent word distributions (topics) that these documents are related to.}
  \label{fig:commits}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
  \caption{Compact-trend-view shows topics per month for MaxDB
    7.500. The x-axis is time in months, the Y axis is used to stack
    topics occurring at the same time. Trends that are continuous are
    plotted as continuous blocks. The top 10 words in the topics are
    joined and embedded in box representing the topics.}
  \label{fig:topicsmear}
\end{figure*}

@@

The \emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to
sink the larger trends to the bottom.  Once the larger trends are
stacked we fill in the gaps with the smaller trends in the same
window, and then stack the smaller trends on top.  Although there is a
chance that gaps will be left due to non-optimal stacking in practice
there are lot of small trends (90\% of all trends contain 1 topic)
that fill in these gaps quickly.  Different instances of the same
trend share the same color; apart from that, the color of trends is
randomly chosen and color similarity is not meaningful.  The
compact-trend-view is convenient because it shows all of the topic
information, as shown by the zoomed-in view of MaxDB 7.500's
compact-trend-view in Figure \ref{fig:zoomedsmear}. This view makes
repeating continuous trends easy to pick out, although discontinuous
trends are harder to spot.

@@

The \emph{trend-histogram} (Figure \ref{fig:histogram}) superficially
resembles the trend-time-line.  However, in this view the trends are
plotted together by stacking to the left of their row, thus time
information is lost.  The trends are ordered by the number of topics
in the trend.  The trend-histogram shows the count of instances of a
trend and thus indicates which trends occur the most. Unfortunately
due to the large number of topics given allotted space it is often
best to crop off the trends with only one topic, otherwise the tail is
long.

@@

%\shrinkit
%\subsection{Datasets and Tools}
%\shrinkit

@@

We applied our methodology to multiple database systems that we
extracted.  To analyze these extracted repositories we used:
Hiraldo-Grok, an OCaml-based variant of the Grok query language;
Gnuplot, a graph plotting package; lda-c, a LDA package implemented by
Blei et al.~\cite{944937}; and our topic plotter, implemented in
Haskell.

@@

When we examined PostgreSQL, we did not find many trends with two or
more topics with a similarity of $7/10$. Those trends that we did find
were not very large, lasting on 3 months at most. The first and second
largest trends were dedicated to the work of two external developers:
Dal Zotto, Dan McGuirk. The fifth largest trend related to work by
PostgreSQL developer D'Arcy. Other topics of the larger trends were
changes to the TODO list in HTML, and time string formating topics
relating to timezones.

If we kept the stop words we found that the large trend consisted
mostly of stop words and non-stop words such as \emph{patch},
\emph{fix}, \emph{update}.  By decreasing the similarity constraint to
$1/2$ similarity the largest most the largest most common trend, which
stretched across development, contained these same words
(\emph{patch}, \emph{fix}, \emph{update}). The second largest trend
mentions Dal Zotto, while the third largest trend mentions the
\texttt{[PATCHES]} mailinglist and the names of some patch
contributors.  Other repeating topics refer to portability with Win32,
Central Europe Time (CEST) from email headers, issues with ALTER TABLE
and CVS branch merging (because CVS does not record merges explicitly).

@@

%XXX Edit
% Was reading lda.report
Bug tracker URLs dominated unique topics during some months. For
instance in the last month of MaxDB 7.500 development every topic
contained 1 unique Bug tracker URL, this pattern did not occur in the
previous month. We investigated the actual revisions and we did find
there was difference in behavior between, and that the last month of
development commits were mostly referencing the bug tracker. If the
topics of one month were 20 or more unique bug tickets being
addressed, the global LDA analysis would probably miss this, yet these
bug tickets were the focus of development for that month and not
necessarily globally relevant.

The query optimizer was a topic that recurred during MaxDB's
development. In our plots, topics that mention \emph{optimizer} occur four
times, yet in the global-trend-view (Figure \ref{fig:statictopics}) it
is not in any of the topics. A query optimizer is rather important to
an DBMS, but as we have shown it just does not pop up as a topic on
its own. We tried to remove words to see if we could get an 
\emph{optimizer} topic. After removing stop words and then two of the most
common words, the global analysis finally had a topic with optimizer in
its topic ten words. This shows that optimizer was important but it
had been obscured, while it would have been noticed using a more
locally relevant approach that we propose.

We noticed that commits that mentioned \emph{PHP} occurred two thirds
less frequently than commits that mentioned \emph{Perl}, but
\emph{Perl}-related topics appeared in the global static topics for
MaxDB while \emph{PHP}-related topics did not.  Our local topic
analysis mentioned \emph{PHP} in 5 different topics, yet only
mentioned \emph{Perl} in 4 different windoed topics and 1 global
topic. Perhaps this is because there was a cluster of \emph{Perl}
mentions during one month while \emph{PHP} was mentions were more
spread out.

We noticed that Perforce and Changelist were in just about every
topic, so we added them to the stop words list. As a result, longer
trends were shortened and in some cases the total number of topics
found per month was less.  Evaluating different similarity thresholds
showed that by removing common words one reduces the general
similarity of topics. That said, the larger topics still clearly
apparent. Thus if more relevant stop words are added one should tune
the topic similarity parameters to handle these changes.

@@

Previous work on topic analysis that employed LSI and LDA typically
extracted a specified number of topics and then tracked them over time. This
means that a single, static set of topics was tracked across the whole
development history of a project. \emph{Global topic analysis} is
topic analysis applied to entire lifetime of a project.

We carried out the same kind of analysis, in order to compare the
results against our time-windowed approach.  To produce
figure~\ref{fig:statictopics}, we extracted 20 topics and plotted the
number of messages per month that were related to that topic. We found
was that one topic would often dominate the results, as shown in the
third row of Figure \ref{fig:statictopics}, while messages related to
other topics did not appear very often.
% This technique seems useful if the 20 extracted topics are relevant
% and useful, that said one weakness of this approach is that documents
% for a topic might only occur in one time window, yet we are tracking
% messages that could be related to it over the entire lifetime of the
% project.
This approach seems reasonable if most of the extracted topics are of
broad interest during most of the development process.  However, it
may be that some topics are of strong interest but only briefly; in
such a case, a window-based model gives a much stronger indication of
the fleeting importance of such topics, and can help to put such a
topic into its proper context.

@@

If we approach the difference of global topic analysis and windowed
topic analysis via common tokens we can see that common token tend to
dominate both results, at least in the case of MaxDB 7.500 our local
topic approach netted us topics which contained these relatively
important and common words: \emph{UNICODE}, \emph{DEC/OSF1},
\emph{ODBC}, \emph{crash}, \emph{SQLCLI}, \emph{SYSDD.cpnix},
\emph{backup}, \emph{select}, \emph{make}, \emph{memory}, \emph{view},
and finally \emph{debug}. \emph{ODBC} is quite important because it
often determines how databases communicate with software. None of
these tokens were part of the global topic analysis topics, but they
were part of 566 commits (6\% of the entire system) to MaxDB
7.500. These tokens were part of 87 out of 520 (26 months, 20 topics
per month) of our locally generated topics.

@@

Even with our liberal topic similarity metrics that produced both long
and short trends, we showed that there are only a few trends in a
repository that recur. Since so few trends recur and so many trends
appear only once this suggests that global topic analysis might be
ignoring locally unique topics.

The utility of global topic analysis is questionable if the value of
information decreases as it becomes older. Perhaps older trends will
dominate the topic analysis. Windowed localized topic analysis shows
what are the unique topics, yet seems to give a more nuanced view of
the timeliness of the important topics.

\begin{figure}[t]
  \centering
  %\includegraphics[width=0.45\textwidth]{maxdb7500-everything-by-month}
  \includegraphics[width=0.45\textwidth]{month}
  \caption{MaxDB 7.500 topics analyzed with global topic analysis, 20
    topics (each row) and their document counts plotted over the
    entire development history of MaxDB 7.500 (26 months). The shade
    of the topic indicates the number of documents matching that topic
    in that month relative to the number of documents (white is most,
    black is least).}
  \label{fig:statictopics}
\end{figure}

@@

In this study we are explicitly trusting that the programmers annotate
their changes with relevant information. We rely on the descriptions
they provide. If the language of check-in comments was automated we
would  be analyzing only that.

We truncated the topics to top 10 tokens, so this summary of a topic
might not have been as useful as determining the actual topic distance
between two word topic distributions. By truncating tokens we could be
losing valuable data.

Each project seemed to have their own set of stop words, and we did not
remove them, but perhaps dropping some of these stop words would aid
the clarity of topics produced through topic analysis; alternatively
it might cause our topic similarity plots to be fundamentally
different.

@@

We proposed and demonstrated the application of windowed topic
analysis, that is, topic analysis applied to commit messages during
periods of development. This approach differs from previous work that
would applied topic analysis globally to the entire history of a
project without regard to time.. We showed that many topics that exist
locally are relevant and interesting yet would often not be detected
via global topic analysis. We identify recurring topics with a topic
similarity measure which looks for topics which recur and mutate
repeatedly throughout the development of the software project.

The value of windowed topic analysis was demonstrated on a case study
of three database systems with a particular focus on MaxDB 7.500. We
showed that global topic analysis missed important topics like such as
ODBC changes.

@@

\shrinkit
\bibliographystyle{latex8}
\bibliography{lda-paper}
